Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-11-29 dsn-2018/bare_conf.tex
More intuition
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

\textbf{TODO - introductory paragraph}

We propose the following metrics for evaluating the utility of a collision
class:
\begin{enumerate}
    \item The number of unique differentials a collision class has.
    \item The number of unit-step neighbors a collision class has.
    \item The maximum count of zeros in a the binary representation of a
        colliding block (and likewise with ones).
    \item Whether there exists a block which collides under multiple initial
        values.
    \item Whether or not zero, one, or both of the blocks in a collision may
        be of ASCII values under any input block difference.
\end{enumerate}
Note that the first three are quantitative measures providing some measure of
flexibility of a collision class, whereas the latter two are merely looking
for a single witness for having the property. Depending on the scenario,
specific properties of a collision may be of more interest than others.

@@

\section{Conclusion} \label{Sec:Conclusion}

The conclusion goes here.

>>>>

\textbf{TODO - audit introductory paragraphs}

We seek to build new intuition about collisions in hash functions,
specifically, collisions in MD4. By developing techniques from logical
cryptanalysis, we seek to complement and extend the results from
differential cryptanalysis. In particular, we wish to use logical cryptanalysis
to discuss the impacts of collisions in real-world systems. In this paper,
we introduce and give justification for general techniques, and present some
results of these techniques.

To begin, real world systems are most vulnerable to two attacks: the preimage
and second preimage. For systems relying on e.g., password validation via
comparing two hashes, preimage attacks would allow inverting the hash,
increasing the risk associated with losing a database of hashed passwords.
On the other hand, systems relying on message validation, such as signature
checks or file integrity checks would be easily broken with fast second
preimage techniques. Both of these attacks are intractable, thus protecting
real-world systems. However, much research has been devoted to collision
attacks, producing efficient results.

Thus, we seek to evaluate the relative difference between a collision
attack and second preimage attacks, to give some measure of distance between
them. Several high profile attacks against hash function usage have been
made, including \textbf{TODO -- cite}, with MD5, cloning a CA certificate
and creating malicious executables, and with SHA-1, a pair of PDFs with the
same hash. However, many of these techniques exploit not the hash, but the
flexibility of the file format, relying instead on embedding arbitrary,
collidable data and later checking for the presence of one of the matching
blocks. This is especially obvious in the MD5 executable and SHA-1 PDF attacks.
In some cases, such as trying to produce a rogue CA certificate that matches
an existing root, this is not possible and thus relies on searching a wide
enough set of objects for a possible match to the collision.

Seeing the need to evaluate how much flexibility exists in a collision,
we propose the following metrics for evaluating the utility of a collision
class:
\begin{enumerate}
    \item The number of unique differentials a collision class has.
    \item The number of unit-step neighbors a collision class has.
    \item The maximum count of zeros in a the binary representation of a
        colliding block (and likewise with ones).
    \item Whether there exists a block which collides under multiple initial
        values.
    \item Whether or not zero, one, or both of the blocks in a collision may
        be of ASCII values under any input block difference.
\end{enumerate}
Note that the first three are quantitative measures providing some measure of
flexibility of a collision class, whereas the latter two are merely looking
for a single witness for having the property. Depending on the scenario,
specific properties of a collision may be of more interest than others.

@@

\section{Conclusion} \label{Sec:Conclusion}
The conclusion goes here.
